Busch, Khakhalin. Intrinsic temporal tuning

### 2019-06-11 merged.tex
Merged text for submission: figure captions to the end

In our previously published large-scale census of tectal cells \citep{ciarleglio2015}, we observed no signs of intrinsic temporal tuning (figures 2, 4), and no interaction between intrinsic and synaptic phenotypes. We argue however, that the current clamp protocols used in earlier studies were not adequate for the task, so in this paper we rely on a dynamic clamp technique. The main benefit of the dynamic clamp, compared to commonly used current step injections, is that it allows a more realistic simulation of neuronal responses to synaptic conductances \citep{prinz2004}. In dynamic clamp, the current injected into the cell is adjusted in real-time, based on a predefined formula that depends on time and cell membrane potential. This means that with dynamic clamp recordings we can excite a neuron in a controlled manner, but still allow both its voltage and transmembrane currents to change, preserving feedback interactions between active currents and transient voltage-gated channel inactivation \citep{ma1996dynamic,zbili2019axonnav}, which is an important mechanism for temporal tuning \citep{branco2010dendritic,platkiewicz2011}. As a second methodological innovation, instead of relying on one type of sensory stimulation \citep{ciarleglio2015}, here we used five different stimulation protocols, and compared their individual effects on intrinsic tuning.

@@

\begin{figure*}
%\includegraphics[width=\linewidth,bb=0 0 7.16in 7.68in]{fig1.pdf}
\includegraphics[width=\linewidth,bb=0 0 516 553]{fig1.pdf}
\caption{
Overview of experimental design and summary of dynamic clamp results. (\textbf{A}). Positions of tectal neurons that were recorded. (\textbf{B}). Sample data from a dynamic clamp experiment. Bottom row: the dynamics of conductances G(t) of four different durations simulated by the dynamic clamp system. Middle row: the currents I(t) dynamically injected into a cell based on conductances of 4 different durations and 3 different amplitudes. Top row: resulting voltage traces V(t) that were recorded and analyzed. (\textbf{C}). A schematic of visual conditioning in “Flash” (left) and “Looming” (right) groups. (\textbf{D}). The number of spikes produced by all neurons in all experiments, split by input peak conductance, and plotted against conductance duration. Black lines show respective averages. (\textbf{E}). A summary of data from D, presented as averages and 95\% confidence intervals.}
\end{figure*}

@@

\begin{figure*}[!t]
%\includegraphics[width=\linewidth,natwidth=9.81in,natheight=6.
\includegraphics[width=\linewidth,bb=0 0 9.81in 6.59in]{fig2.pdf}
\caption{
Quantification of changes in temporal tuning in response to sensory experience. (\textbf{A}). An illustration of how “Temporal tuning” and “Amplitude tuning” values were calculated. For the temporal tuning measure, the value of zero corresponds to linear dependency (blue line), positive values - to an accelerating, supralinear curve (red), and negative values - to plateau- and hill-shaped curves (gray, orange, green). For amplitude tuning, higher values correspond to faster increase in spiking with increased conductance. (\textbf{B}). Amplitude tuning of neurons across different experimental groups (stars show t-test p$<$0.05 compared to control). (\textbf{C}). Temporal tuning and spikiness of neurons in different experimental groups. Neurons from each target group are shown in color, while all neurons from all groups are shown in light gray, as a reference; groups means are shown as black dots; ellipses represent 95\% normal confidence regions. Two outliers (top right corner) are brought within the axes limits. (\textbf{D}). Same data as in (C), shown as averages for each group, with 95\% confidence intervals. Black arrows show the effects of sound clicks, when they were added to control (red), and when they were added to "Flashes" (brown), to form two types of multisensory stimuli.}
\end{figure*}

@@

\begin{figure*}[!t]
%\includegraphics[width=\linewidth,natwidth=8.83in,natheight=5.02in]{fig3.pdf}
\includegraphics[width=\linewidth,bb=0 0 8.83in 5.02in]{fig3.pdf}
\caption{Changes in synaptic transmission, and co-tuning of synaptic and intrinsic neuronal properties. (\textbf{A}). Amplitudes of early monosynaptic inputs to tectal neurons in all experimental groups (in log scale, outliers brought within the axes limits, stars show t-test p$<$0.05 compared to control). (\textbf{B}). A sample synaptic recording, showing all traces for one cell (green) and an average trace (blue). The black bars show the areas at which early monosynaptic and late polysynaptic currents were measured; the vertical position of each bar represents the respective average current. The second, longer bar does not completely fit within the figure at this scale. (\textbf{C}). Synaptic current duration (vertical axis) was mostly defined by the amplitude of early monosynaptic inputs (horizontal axis). (\textbf{D}). Synaptic current durations were different between experimental groups (see text; stars show t-test p$<$0.05 compared to control). (\textbf{E}). Across experiments, average temporal tuning in each group positively correlated with the average durations of synaptic currents they received. (\textbf{F}). Within experimental groups, temporal tuning of individual neurons negatively correlated with the duration of synaptic currents they received. Axes show within-group deviations of temporal tuning and synaptic current duration from respective averages for each group.}
\end{figure*}

@@

\begin{figure}[!t]
\centering
%\includegraphics[width=3.2in,natwidth=4.57in,natheight=5.71in]{fig4.pdf}
\includegraphics[width=3.2in,bb=0 0 4.57in 5.71in]{fig4.pdf}
\caption{Electrophysiological properties of individual neurons compared to their spiking in current and dynamic clamp experiments. (\textbf{A}). A set of curves from one voltage step experiment; black bars show the areas used to average Na (bottom) and transient K (top) currents (see Methods). (\textbf{B}). Processing of ionic currents data with IV-curves translated into two parameters (threshold potential and peak current) for each ionic conductance. (\textbf{C}). Sample data from one current clamp experiment; spikes are marked with black dots. (\textbf{D}). Estimations of excitability from current clamp experiments (horizontal axis) and dynamic clamp experiments (vertical axis) correlate. (\textbf{E}). The number of spikes registered in current clamp mode: values predicted from a linear model plotted against observed values. The model works reasonably well (61\% of variance explained). (\textbf{F}). Similar comparison for the dynamic clamp experiments: the model has very low predictive value (13\% of variance explained). Here and in E, both values are adjusted for position.}
\end{figure}

@@

>>>>

In our previously published large-scale census of tectal cells \citep{ciarleglio2015}, we observed no signs of intrinsic temporal tuning (Figures 2, 4), and no interaction between intrinsic and synaptic phenotypes. We argue however, that the current clamp protocols used in earlier studies were not adequate for the task, so in this paper we rely on a dynamic clamp technique. The main benefit of the dynamic clamp, compared to commonly used current step injections, is that it allows a more realistic simulation of neuronal responses to synaptic conductances \citep{prinz2004}. In dynamic clamp, the current injected into the cell is adjusted in real-time, based on a predefined formula that depends on time and cell membrane potential. This means that with dynamic clamp recordings we can excite a neuron in a controlled manner, but still allow both its voltage and transmembrane currents to change, preserving feedback interactions between active currents and transient voltage-gated channel inactivation \citep{ma1996dynamic,zbili2019axonnav}, which is an important mechanism for temporal tuning \citep{branco2010dendritic,platkiewicz2011}. As a second methodological innovation, instead of relying on one type of sensory stimulation \citep{ciarleglio2015}, here we used five different stimulation protocols, and compared their individual effects on intrinsic tuning.

@@

@@

@@

@@

@@

\subsection*{Figure captions}

\textbf{Figure 1}

Overview of experimental design and summary of dynamic clamp results. (\textbf{A}). Positions of tectal neurons that were recorded. (\textbf{B}). Sample data from a dynamic clamp experiment. Bottom row: the dynamics of conductances G(t) of four different durations simulated by the dynamic clamp system. Middle row: the currents I(t) dynamically injected into a cell based on conductances of 4 different durations and 3 different amplitudes. Top row: resulting voltage traces V(t) that were recorded and analyzed. (\textbf{C}). A schematic of visual conditioning in “Flash” (left) and “Looming” (right) groups. (\textbf{D}). The number of spikes produced by all neurons in all experiments, split by input peak conductance, and plotted against conductance duration. Black lines show respective averages. (\textbf{E}). A summary of data from D, presented as averages and 95\% confidence intervals

\textbf{Figure 2}

Quantification of changes in temporal tuning in response to sensory experience. (\textbf{A}). An illustration of how “Temporal tuning” and “Amplitude tuning” values were calculated. For the temporal tuning measure, the value of zero corresponds to linear dependency (blue line), positive values - to an accelerating, supralinear curve (red), and negative values - to plateau- and hill-shaped curves (gray, orange, green). For amplitude tuning, higher values correspond to faster increase in spiking with increased conductance. (\textbf{B}). Amplitude tuning of neurons across different experimental groups (stars show t-test p$<$0.05 compared to control). (\textbf{C}). Temporal tuning and spikiness of neurons in different experimental groups. Neurons from each target group are shown in color, while all neurons from all groups are shown in light gray, as a reference; groups means are shown as black dots; ellipses represent 95\% normal confidence regions. Two outliers (top right corner) are brought within the axes limits. (\textbf{D}). Same data as in (C), shown as averages for each group, with 95\% confidence intervals. Black arrows show the effects of sound clicks, when they were added to control (red), and when they were added to "Flashes" (brown), to form two types of multisensory stimuli.

\textbf{Figure 3}

Changes in synaptic transmission, and co-tuning of synaptic and intrinsic neuronal properties. (\textbf{A}). Amplitudes of early monosynaptic inputs to tectal neurons in all experimental groups (in log scale, outliers brought within the axes limits, stars show t-test p$<$0.05 compared to control). (\textbf{B}). A sample synaptic recording, showing all traces for one cell (green) and an average trace (blue). The black bars show the areas at which early monosynaptic and late polysynaptic currents were measured; the vertical position of each bar represents the respective average current. The second, longer bar does not completely fit within the figure at this scale. (\textbf{C}). Synaptic current duration (vertical axis) was mostly defined by the amplitude of early monosynaptic inputs (horizontal axis). (\textbf{D}). Synaptic current durations were different between experimental groups (see text; stars show t-test p$<$0.05 compared to control). (\textbf{E}). Across experiments, average temporal tuning in each group positively correlated with the average durations of synaptic currents they received. (\textbf{F}). Within experimental groups, temporal tuning of individual neurons negatively correlated with the duration of synaptic currents they received. Axes show within-group deviations of temporal tuning and synaptic current duration from respective averages for each group.

\textbf{Figure 4}

Electrophysiological properties of individual neurons compared to their spiking in current and dynamic clamp experiments. (\textbf{A}). A set of curves from one voltage step experiment; black bars show the areas used to average Na (bottom) and transient K (top) currents (see Methods). (\textbf{B}). Processing of ionic currents data with IV-curves translated into two parameters (threshold potential and peak current) for each ionic conductance. (\textbf{C}). Sample data from one current clamp experiment; spikes are marked with black dots. (\textbf{D}). Estimations of excitability from current clamp experiments (horizontal axis) and dynamic clamp experiments (vertical axis) correlate. (\textbf{E}). The number of spikes registered in current clamp mode: values predicted from a linear model plotted against observed values. The model works reasonably well (61\% of variance explained). (\textbf{F}). Similar comparison for the dynamic clamp experiments: the model has very low predictive value (13\% of variance explained). Here and in E, both values are adjusted for position.
